Food-web complexity flattens the fitness landscape of an insect herbivore

### 2018-06-21 elsevier_test/elsevier_test.Rmd
added GRC abstract to think about reframing the paper.
Former-commit-id: f447ab24724625427d952dae9067aee93b395039

---
title: Food-web complexity flattens the fitness landscape of an insect herbivore
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich, Switzerland
    footnote: Corresponding Author
  - name: Jordi Bascompte
    email: jordi.bascompte@ieu.uzh.ch
    affiliation: Another University
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: |
  Studies of natural selection and fitness landscapes usually treat the network of interacting species as a “black box”. Given that the loss of biodiversity is simplifying the structure of ecological networks, there is a pressing need to answer the question: how does network complexity affect natural selection and the fitness landscape of associated species? To answer this question, we conducted a field experiment that manipulated the complexity of a food web associated with a galling insect herbivore. To maintain complex food webs, we allowed the entire community of natural enemies to attack insect galls on 64 plants in a common garden setting. To create simple food webs, we excluded a guild of three larval parasitoids by bagging galls on 64 different plants; therefore, mortality in this treatment was primarily due to a single egg parasitoid that attacks prior to gall formation. We then measured herbivore survival as a function of three key gall traits in each treatment. We found that more traits were under selection in the simple vs. complex food web. This occurred because different parasitoid species impose different selection pressures on gall traits, thereby minimizing relative fitness differences among insect galls with different phenotypes. Our work suggests that more complex food webs allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution to environmental change.

>>>>

---
title: Food-web complexity flattens the fitness landscape of an insect herbivore
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: Department of Evolutionary Biology and Environmental Studies, University of Zurich, Zurich, Switzerland
    footnote: Corresponding Author
  - name: Jordi Bascompte
    email: jordi.bascompte@ieu.uzh.ch
    affiliation: Another University
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: Studies of natural selection and fitness landscapes usually treat the network of interacting species as a “black box”. Given that the loss of biodiversity is simplifying the structure of ecological networks, there is a pressing need to answer the question: how does network complexity affect natural selection and the fitness landscape of associated species? To answer this question, we conducted a field experiment that manipulated the complexity of a food web associated with a galling insect herbivore. To maintain complex food webs, we allowed the entire community of natural enemies to attack insect galls on 64 plants in a common garden setting. To create simple food webs, we excluded a guild of three larval parasitoids by bagging galls on 64 different plants; therefore, mortality in this treatment was primarily due to a single egg parasitoid that attacks prior to gall formation. We then measured herbivore survival as a function of three key gall traits in each treatment. We found that more traits were under selection in the simple vs. complex food web. This occurred because different parasitoid species impose different selection pressures on gall traits, thereby minimizing relative fitness differences among insect galls with different phenotypes. Our work suggests that more complex food webs allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution to environmental change.

Species-interaction networks provide a mechanistic link between community and population ecology, as they describe who interacts with whom in a community. Similarly, fitness landscapes provide a mechanistic link between population and evolutionary ecology, as they describe the influence of natural selection on phenotypic variation. It remains unclear, however, how the network structure of species interactions shapes the fitness landscape. Understanding this relationship is likely key for developing a predictive ecology across scales of biological organization. To examine the relationship between network structure and the fitness landscape, we conducted a field experiment that manipulated the network of trophic interactions (simple vs. complex) influencing the fitness of an insect herbivore. We then quantified the fitness landscape of herbivores in each treatment by measuring herbivore survival as a function of multiple phenotypic traits. We found that more traits were under selection in the simple vs. complex trophic network. This occurred because different natural enemies impose different selection pressures on herbivore traits, thereby minimizing relative fitness differences among herbivore phenotypes in the complex network. Our work suggests that more complex trophic networks allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution of populations to environmental change.
